---
output:
  beamer_presentation:
    theme: "Frankfurt"
    colortheme: "orchid"
    slide_level: 3
    includes:
      in_header:
        beamer_template.txt
    keep_tex: true
classoption: aspectratio=169
editor_options: 
  chunk_output_type: console
---
\title{Attraction Effect in Risky Choices}
\author{Yi-tsen Liao}
\institute{Carnegie Mellon University Social Decision Science}
\date{Sep 29, 2025\\BDR Lab Meeting}

\maketitle

# Introduction
```{r setup, include = F}
library(magrittr)
set.seed(2000)

knitr::opts_chunk$set(echo = F, message = F, warning = F,
                      fig.width = 5.5, fig.height = 2.5,
                      fig.align = "right")
  
study1 <- readRDS(here::here("data/study1.Rds"))
study1_complete <- readRDS(here::here("data/study1_complete.Rds"))
study1_change <- readRDS(here::here("data/study1_change.Rds"))


```

## Intro
### Context Effect
\begin{itemize}
  \item When choosing among options with multiple attributes
  \pause
  \item We often face difficult situations where one option is better in some domains,
  \pause
  \item while the alternative is better in the other domains
\end{itemize}

\vfill

\begin{center}
\pause
  \textbf{Choosing partner, houses hunting, buying cars...}
\end{center}


### Context Effect
\begin{center}
  \textbf{These choices may be influenced by other available options}
\end{center}

\pause
\begin{itemize}
  \item An option may be viewed differently when it is evaluated in isolation v.s. when it is jointly evaluated with other options (Hsee et al, 1999)
  \pause
  \item When making choices among options with multiple "aspects", we often compare them attribute by attribute (Tversky, 1972)
  \pause
  \item A choice between two options is affected by the
introduction of a third option (Trueblood et al, 2014)
\end{itemize}



### Context Effect
\begin{center}
  \textbf{Types of context effects vary by how the third option is constructed}
\end{center}

\pause
\begin{itemize}
  \item Attraction effect (Huber et al, 1982)
  \begin{itemize}
    \item The third option ("Decoy") is similar but inferior to the focal option 
    \item which enhances the probability of choosing the "similar but better" focal option
  \end{itemize}
  \pause
  \item Similarity effect (Tversky, 1972)
  \begin{itemize}
    \item The third option is similar to the non-focal option and DM is indifferent to all three options
    \item which enhances the probability of choosing the "unique" focal option
  \end{itemize}
  \pause
  \item Compromise effect (Simonson, 1989)
  \begin{itemize}
    \item The third is an extreme alternative to the focal option and DM is indifferent to all three options
    \item which enhances the probability of choosing the "moderate" focal option
  \end{itemize}
\end{itemize}

\vfill

\begin{center}
\pause
  \textbf{In this study, we focus on the attraction effect!}
\end{center}


### Attraction Effect
\begin{center}
  \textbf{Types of attraction effects also vary by how the "decoy" is constructed}
\end{center}

\pause
\begin{itemize}
  \item Frequency decoy
  \begin{itemize}
    \item The decoy is worse than the focal option in its \textbf{weak} attribute
  \end{itemize}
  \pause
  \item Range decoy
  \begin{itemize}
    \item The decoy is worse than the focal option in its \textbf{strong} attribute
  \end{itemize}
  \pause
  \item Range-Frequency decoy
  \begin{itemize}
    \item The decoy is worse than the focal option in \textbf{all} attributes
  \end{itemize}
\end{itemize}
  
### Literatures on Attraction Effect
\begin{itemize}
  \pause
  \item Less studies on the mechanism of how the attraction effect
  \pause
  \item Often compare between focal and non-focal option, instead of a control
  \pause
  \item Often assume these different types of decoys work in the same way
  \pause
  \item Different theories may predict different choices given the same choice set
\end{itemize}

### Application of Attraction Effect
\begin{center}
  \textbf{In pricing}
\end{center}
\begin{center}
  \includegraphics[width=0.6\textwidth]{intro_pricing.png}
\end{center}

\vfill
\hfill \tiny Source: @UXMental https://www.instagram.com/p/CsG2uBCNxO6/

### Application of Attraction Effect
\begin{center}
  \textbf{In how we evaluate "beauty"}
\end{center}
\begin{center}
  \includegraphics[width=0.6\textwidth]{intro_beauty.png}
\end{center}

\vfill
\hfill \tiny Source: @Dan Ariely https://www.youtube.com/watch?v=9X68dm92HVI

### Application of Attraction Effect
\begin{center}
  \textbf{In visual perception...}
\end{center}
\begin{center}
  \includegraphics[width=0.6\textwidth]{intro_perception.png}
\end{center}

\vfill
\hfill \tiny Trueblood, J. S., Brown, S. D., Heathcote, A., & Busemeyer, J. R. (2013). Not just for consumers: Context effects are fundamental to decision making. Psychological science, 24(6), 901-908.

### Limits of Studies on Attraction Effect
\begin{itemize}
  \pause
  \item Less studies on the mechanism of how attraction effect works exactly
  \pause
  \item Often compare between focal and non-focal option, instead of a control
  \pause
  \item Often assume these different types of decoys work in the same way
  \pause
  \item Different theories may predict different choices given the same choice set
\end{itemize}

### Limits of Studies on Attraction Effect
\begin{itemize}
  \item Has been widely studied in various fields, with real-world application
  \pause
  \item Caveat: they may not always work (Frederick et al, 2014)
  \pause
  \item Restricted to very limited number of options (most commonly 3)
  \pause
  \item with very limited number of attributes (most commonly 2)
\end{itemize}


\vfill
\begin{center}
  \textbf{What are the settings that are simplistic enough to fit into these narrow criteria?}
\end{center}

\begin{center}
  \pause
  \textbf{Good old lottery choices :)}
\end{center}

### Risky Choices
\begin{itemize}
  \item People have a hard time thinking about risk
  \pause
  \item And we have a long history studying the inconsistency in choices under uncertainty
  \pause
  \item A lottery in some sense is an option with only two attributes: 
  \pause
  \begin{itemize}
    \item A set of prizes
    \item And a set of probabilities associated with the prizes
  \end{itemize}
\end{itemize}


\vfill
\begin{center}
\pause
  \textbf{Choices among lotteries are essentially tradeoffs between these two attributes}
\end{center}

# Design
### Simple Two-state No-Lose Lotteries

\begin{itemize}
  \item Positive payoff in one state, nothing in the other
  \pause
  \item Choose among three lotteries: Focal, Non-focal and Decoy
  \begin{itemize}
    \pause
    \item Focal and Non-focal lotteries are symmetrical
    \pause
    \item Treatment Decoy: same winning probability, 1/2 of the payoff
    \pause
    \item Control Decoy has 50\% of winning probability, same expected payoff as the Treatment
  \end{itemize}
  \pause
  \item Decisions are incentivized!
  
\end{itemize}

\vfill
\hfill \tiny *All studies preregistered on AsPredicted


### Study 1: Control Trails
\begin{center}
  \includegraphics[width=1\textwidth]{Control.png}
\end{center}

### Study 1: Treatment Trails: targets Safe lottery
\begin{center}
  \includegraphics[width=1\textwidth]{Treatment_Risk_Averse.png}
\end{center}

### Study 1: Treatment Trails: targets Risky lottery
\begin{center}
  \includegraphics[width=1\textwidth]{Treatment_Risk_Seeking.png}
\end{center}

### Study 1: Design
\begin{itemize}
  \item 600 participants from Prolific
  \pause
  \item Within subject treatment
  \begin{itemize}
    \pause
    \item 10 trails of lottery decision
    \item 5 unique pairs of Focal and Non-focal lotteries
    \item Along with either a Treatment or a Control Decoy
  \end{itemize}
  \pause
  \item Between subject condition
  \begin{itemize}
    \pause
    \item \textbf{Risky} Treatment Decoy targets the risky lottery
    \item \textbf{Safe} Treatment Decoy targets the safe lottery
  \end{itemize}
  \pause
  \item Exclude participants who choose the decoy more than 5 times
\end{itemize}

\vfill
\hfill \tiny *All studies preregistered on AsPredicted

### Different levels of tradeoffs
\begin{itemize}
  \item 95\% of 20¢; 20\% of 95¢
  \item 90\% of 25¢; 25\% of 90¢
  \item 85\% of 30¢; 30\% of 85¢
  \item 80\% of 35¢; 35\% of 80¢
  \item 75\% of 40¢; 40\% of 75¢
\end{itemize}

### Study 1: Hypothesis
\begin{itemize}
  \item Can the Treatment Decoy increase the likelihood of choosing the Focal lottery?
  \pause
  \item Does the effect depend on specific characteristic in lottery pairs?
  \pause
  \item Whether Treatment Decoy is better at enhancing risky lotteries or safe lotteries?
\end{itemize}

# Results

### Study 1: Seperated by Conditions
```{r}
study1 %>%
  dplyr::mutate(condition = dplyr::recode_factor(condition,
                                                             'risk_averse' = 'Safe condition',
                                                             'risk_seeking' = 'Risky condition'),
                treatment = dplyr::recode_factor(treatment,
                                                             'control' = 'Control decoy',
                                                             'treatment' = 'Treatment decoy')) %>%
  dplyr::group_by(condition, treatment) %>%
  dplyr::summarize(mean = mean(choice),
                   n = dplyr::n(),
                   sd = sd(choice),
                   ci = qnorm(0.975) * sd/sqrt(n)) %>%
  ggplot2::ggplot(ggplot2::aes(x = condition,
                               y = mean,
                               fill = treatment)) +
  ggplot2::geom_col(position = 'dodge') +
  ggplot2::geom_text(ggplot2::aes(x = condition,
                                  y = 0.05,
                                  label = paste0(round(mean*100, 2), "%", "\n",
                                                 "n = ", n)),
                     color = 'white',
                     lineheight = 0.75,
                     size = 2.5,
                     position = ggplot2::position_dodge(0.9)) +
  ggplot2::coord_cartesian(ylim = c(0, 1),
                           expand = 0) +
  ggplot2::scale_fill_manual(values = c('#998ec3', '#f1a340')) +
  ggplot2::geom_errorbar(ggplot2::aes(ymin = mean - ci,
                                      ymax = mean + ci),
                         position = ggplot2::position_dodge(0.9),
                         width = 0.2) +
  papaja::theme_apa() +
  ggplot2::theme(legend.position = c(1, 1.1),
                 legend.justification = c(1, 1),
                 legend.direction = "horizontal",
                 legend.background = ggplot2::element_rect(fill = "transparent", color = NA),
                 legend.text = ggplot2::element_text(size = 6),
                 legend.title = ggplot2::element_blank(),
                 axis.title.x = ggplot2::element_blank(),
                 axis.title.y = ggplot2::element_blank(),
                 axis.ticks.x = ggplot2::element_blank(),
                 strip.placement = "outside")+
  ggplot2::scale_y_continuous("Proportion Choosing Focal Lottery")
```

### Study 1: No and Opposite Preference Reversal (Treatment)

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(clubSandwich)
library(broom)

# Compute participant-level means
clustered_estimates <- study1_complete %>%
  filter(treatment == "treatment") %>%
  mutate(
    choose_risky = ifelse(raw_choice == "risky", 1, 0),
    choose_safe  = ifelse(raw_choice == "safe", 1, 0),
    choose_decoy  = ifelse(raw_choice == "decoy", 1, 0)
  ) %>%
  group_by(condition, PROLIFIC_PID) %>%
  summarise(
    risky_rate = mean(choose_risky),
    safe_rate = mean(choose_safe),
    decoy_rate = mean(choose_decoy),
    .groups = "drop"
  )

# Ensure condition is a factor with reference = "risk_averse"
clustered_estimates <- clustered_estimates %>%
  mutate(condition = factor(condition, levels = c("risk_averse", "risk_seeking")))

# Fit linear models
mod_risky <- lm(risky_rate ~ condition, data = clustered_estimates)
mod_safe  <- lm(safe_rate  ~ condition, data = clustered_estimates)
mod_decoy  <- lm(decoy_rate  ~ condition, data = clustered_estimates)

# Clustered SEs
risky_cl <- coef_test(mod_risky, vcov = "CR2", cluster = clustered_estimates$PROLIFIC_PID)
safe_cl  <- coef_test(mod_safe,  vcov = "CR2", cluster = clustered_estimates$PROLIFIC_PID)
decoy_cl  <- coef_test(mod_decoy,  vcov = "CR2", cluster = clustered_estimates$PROLIFIC_PID)

# Inspect result tables
print(risky_cl)
print(safe_cl)
print(decoy_cl)

# Helper function to safely extract coefficient and SE
get_coef_info <- function(tbl) {
  tbl %>%
    dplyr::filter(Coef != "(Intercept)") %>%
    dplyr::summarise(
      diff = beta,
      se   = SE
    ) %>%
    as.list()
}

# Extract risky and safe info
risky_info <- get_coef_info(risky_cl)
safe_info  <- get_coef_info(safe_cl)
decoy_info  <- get_coef_info(decoy_cl)

# Combine into preference reversal table
pref_rev <- tibble(
  lottery = c("Risky lottery", "Safe lottery","Decoy"),
  diff = c(risky_info$diff, -safe_info$diff, decoy_info$diff),  # flip sign for safe
  se = c(risky_info$se, safe_info$se, decoy_info$se)
) %>%
  mutate(
    ci_low = diff - 1.96 * se,
    ci_high = diff + 1.96 * se
  )

Preference_Reversal_P<-ggplot(pref_rev, aes(x = lottery, y = diff)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high),
                width = 0.15, linewidth = 0.6, color = "gray30") +
  geom_point(size = 4, color = "black") +
  geom_line(aes(group = 1), color = "gray50", linewidth = 0.8) +
  coord_cartesian(ylim = c(-0.13, 0.13), expand = 0) +
  labs(
    x = "",
    y = "P(Focal) - P(Nonfocal)",
    title = "Treatment Risky Decoy v.s. Treatment Safe Decoy",
    size=20,
    face="bold"
    
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold")
  )
```
```{r, fig.width=5.5, fig.height=3}
Preference_Reversal_P
```

### No Heterogenous Effect Across Pairs
```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(clubSandwich)
library(broom)

# Step 1. Compute participant-level means per pair
clustered_estimates_pairs <- study1_complete %>%
  filter(treatment == "treatment") %>%
  mutate(
    choose_risky = ifelse(raw_choice == "risky", 1, 0),
    choose_safe  = ifelse(raw_choice == "safe", 1, 0)
  ) %>%
  group_by(condition, PROLIFIC_PID, pair) %>%  # pair_id = unique trial set
  summarise(
    risky_rate = mean(choose_risky),
    safe_rate  = mean(choose_safe),
    .groups = "drop"
  ) %>%
  mutate(condition = factor(condition, levels = c("risk_averse", "risk_seeking")))

# Step 2. Define helper to fit models with clustered SEs per pair
get_clustered_diff <- function(df, outcome) {
  mod <- lm(reformulate("condition", outcome), data = df)
  cl  <- coef_test(mod, vcov = "CR2", cluster = df$PROLIFIC_PID)
  cl_df <- as.data.frame(cl)
  est <- cl_df |> dplyr::filter(Coef == "conditionrisk_seeking")
  tibble(
    diff = est$beta,
    se = est$SE
  )
}

# Step 3. Loop through each pair to extract both risky & safe diffs
pair_results <- clustered_estimates_pairs %>%
  group_by(pair) %>%
  summarise(
    risky_diff = get_clustered_diff(cur_data(), "risky_rate")$diff,
    risky_se   = get_clustered_diff(cur_data(), "risky_rate")$se,
    safe_diff  = get_clustered_diff(cur_data(), "safe_rate")$diff,
    safe_se    = get_clustered_diff(cur_data(), "safe_rate")$se,
    .groups = "drop"
  )

# Step 4. Reshape and flip safe sign for consistency
pref_rev_pairs <- pair_results %>%
  mutate(safe_diff = -safe_diff) %>%
  tidyr::pivot_longer(
    cols = c(risky_diff, safe_diff, risky_se, safe_se),
    names_to = c("lottery", ".value"),
    names_pattern = "(risky|safe)_(diff|se)"
  ) %>%
  mutate(
    lottery = factor(lottery, levels = c("risky", "safe"),
                     labels = c("Risky lottery", "Safe lottery")),
    ci_low = diff - 1.96 * se,
    ci_high = diff + 1.96 * se
  )

# Step 5. Plot: side-by-side effects by pair
Hetero_P<-ggplot(pref_rev_pairs, aes(x = factor(pair), y = diff, color = lottery, group = lottery)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high),
                position = position_dodge(width = 0.4), width = 0.15) +
  geom_point(position = position_dodge(width = 0.4), size = 3) +
  geom_line(position = position_dodge(width = 0.4)) +
  coord_cartesian(ylim = c(-0.15, 0.15), expand = 0) +
  scale_color_manual(values = c("#3E8EDE", "#E69F00")) +
  labs(
    x = "Lottery pair",
    y = "P(Focal) - P(Nonfocal)",
    title = "Treatment Risky Decoy v.s. Treatment Safe Decoy",
    color = "Lottery type",
  ) +
  theme_minimal(base_size = 10) +
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(size = 11)
  )
```

```{r, fig.width=5.5, fig.height=3}
Hetero_P
```


### Study 1: Risky Decoy: Treatment v.s. Control
```{r,include=FALSE}
clustered_estimates1 <- study1_complete %>%
  filter(condition == "risk_seeking") %>%
  mutate(
    choose_risky = ifelse(raw_choice == "risky", 1, 0),
    choose_safe  = ifelse(raw_choice == "safe", 1, 0),
    choose_decoy = ifelse(raw_choice == "decoy", 1, 0)
  ) %>%
  group_by(treatment, PROLIFIC_PID) %>%
  summarise(
    risky_rate = mean(choose_risky),
    safe_rate = mean(choose_safe),
    decoy_rate = mean(choose_decoy),
    .groups = "drop"
  )

# Ensure condition is a factor with reference = "risk_averse"
clustered_estimates1 <- clustered_estimates1 %>%
  mutate(treatment = factor(treatment, levels = c("control", "treatment")))

# Fit linear models
mod_risky1 <- lm(risky_rate ~ treatment, data = clustered_estimates1)
mod_safe1  <- lm(safe_rate  ~ treatment, data = clustered_estimates1)
mod_decoy1  <- lm(decoy_rate  ~ treatment, data = clustered_estimates1)


# Clustered SEs
risky_cl1 <- coef_test(mod_risky1, vcov = "CR2", cluster = clustered_estimates1$PROLIFIC_PID)
safe_cl1  <- coef_test(mod_safe1,  vcov = "CR2", cluster = clustered_estimates1$PROLIFIC_PID)
decoy_cl1  <- coef_test(mod_decoy1,  vcov = "CR2", cluster = clustered_estimates1$PROLIFIC_PID)

# Inspect result tables
print(risky_cl1)
print(safe_cl1)
print(decoy_cl1)

# Helper function to safely extract coefficient and SE
get_coef_info <- function(tbl) {
  tbl %>%
    dplyr::filter(Coef != "(Intercept)") %>%
    dplyr::summarise(
      diff = beta,
      se   = SE
    ) %>%
    as.list()
}

# Extract risky and safe info
risky_info1 <- get_coef_info(risky_cl1)
safe_info1  <- get_coef_info(safe_cl1)
decoy_info1  <- get_coef_info(decoy_cl1)

# Combine into preference reversal table
pref_rev1 <- tibble(
  lottery = c("Risky lottery", "Safe lottery", "Decoy"),
  diff = c(risky_info1$diff, safe_info1$diff, decoy_info1$diff),  # flip sign for safe
  se = c(risky_info1$se, safe_info1$se, decoy_info1$se)
) %>%
  mutate(
    ci_low = diff - 1.96 * se,
    ci_high = diff + 1.96 * se
  )

p_risk_seeking<-ggplot(pref_rev1, aes(x = lottery, y = diff)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high),
                width = 0.15, linewidth = 0.6, color = "gray30") +
  geom_point(size = 4, color = "black") +
  geom_line(aes(group = 1), color = "gray50", linewidth = 0.8) +
  coord_cartesian(ylim = c(-0.15, 0.15), expand = 0) +
  labs(
    x = "",
    y = "P(Treatment) - P(Control)",
    title = "Treatment Risky Decoy v.s. Control Decoy",
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold")
  )


```
```{r, fig.width=5.5, fig.height=3}
p_risk_seeking
```
### Study 1: Safe Decoy: Treatment vs. Control
```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(clubSandwich)
library(broom)

# Risk_Averse Condition

# Compute participant-level means
clustered_estimates <- study1_complete %>%
  filter(condition == "risk_averse") %>%
  mutate(
    choose_risky = ifelse(raw_choice == "risky", 1, 0),
    choose_safe  = ifelse(raw_choice == "safe", 1, 0),
    choose_decoy = ifelse(raw_choice == "decoy", 1, 0)
  ) %>%
  group_by(treatment, PROLIFIC_PID) %>%
  summarise(
    risky_rate = mean(choose_risky),
    safe_rate = mean(choose_safe),
    decoy_rate = mean(choose_decoy),
    .groups = "drop"
  )

# Ensure condition is a factor with reference = "risk_averse"
clustered_estimates <- clustered_estimates %>%
  mutate(treatment = factor(treatment, levels = c("control", "treatment")))

# Fit linear models
mod_risky <- lm(risky_rate ~ treatment, data = clustered_estimates)
mod_safe  <- lm(safe_rate  ~ treatment, data = clustered_estimates)
mod_decoy  <- lm(decoy_rate  ~ treatment, data = clustered_estimates)


# Clustered SEs
risky_cl <- coef_test(mod_risky, vcov = "CR2", cluster = clustered_estimates$PROLIFIC_PID)
safe_cl  <- coef_test(mod_safe,  vcov = "CR2", cluster = clustered_estimates$PROLIFIC_PID)
decoy_cl  <- coef_test(mod_decoy,  vcov = "CR2", cluster = clustered_estimates$PROLIFIC_PID)

# Inspect result tables
print(risky_cl)
print(safe_cl)
print(decoy_cl)

# Helper function to safely extract coefficient and SE
get_coef_info <- function(tbl) {
  tbl %>%
    dplyr::filter(Coef != "(Intercept)") %>%
    dplyr::summarise(
      diff = beta,
      se   = SE
    ) %>%
    as.list()
}

# Extract risky and safe info
risky_info <- get_coef_info(risky_cl)
safe_info  <- get_coef_info(safe_cl)
decoy_info  <- get_coef_info(decoy_cl)

# Combine into preference reversal table
pref_rev <- tibble(
  lottery = c("Risky lottery", "Safe lottery", "Decoy"),
  diff = c(risky_info$diff, safe_info$diff, decoy_info$diff),  # flip sign for safe
  se = c(risky_info$se, safe_info$se, decoy_info$se)
) %>%
  mutate(
    ci_low = diff - 1.96 * se,
    ci_high = diff + 1.96 * se
  )

p_risk_averse<-ggplot(pref_rev, aes(x = lottery, y = diff)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high),
                width = 0.15, linewidth = 0.6, color = "gray30") +
  geom_point(size = 4, color = "black") +
  geom_line(aes(group = 1), color = "gray50", linewidth = 0.8) +
  coord_cartesian(ylim = c(-0.15, 0.15), expand = 0) +
  labs(
    x = "",
    y = "P(Treatment) - P(Control)",
    title = "Treatment Safe Decoy v.s. Control Decoy",
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold")
  )
```
```{r, fig.width=5.5, fig.height=3}
p_risk_averse
```


### Study 1: Conclusion
\begin{flushleft}
  \textbf{No significant preference reversal}
  \begin{itemize}
    \item Compared with risky decoy, Safe decoy slightly \textit{reduced} safe choice
    \item AE not robust to visual representation (Frederick et al., 2014)
  \end{itemize}
\end{flushleft}

\pause
\begin{flushleft}
  \textbf{No heterogenous effect across pairs characteristics}
  \begin{itemize}
    \item Distance between competing options didn't matter
    \item No support for similar effect on perceptual task (Farmer et al., 2016)
    \item Similarity doesn't make the preference easier shift
  \end{itemize}
\end{flushleft}

\pause
\begin{flushleft}
  \textbf{Control decoy is "inadvertently" dominated by safe lottery on both attributes}
  \begin{itemize}
    \item Control decoy \textit{boosted} safe choices more than safe decoy
    \item Must decoys be strictly dominated in \textit{both} attributes?
    \item Couldn't similarity itself trigger comparison?
  \end{itemize}
\end{flushleft}

### Study 1: Make One Attractive or Avoid Worst Option?
\begin{flushleft}
  \textbf{47\% \$23 v.s. 19\% \$57}
  \begin{itemize}
    \item Hard to tell which has higher EV
    \item Decoy 19\% \$28 may work
    \item At least I won't choose the worst option
  \end{itemize}
\end{flushleft}

\pause
\begin{flushleft}
  \textbf{47\% \$23 v.s. 23\% \$47}
  \begin{itemize}
    \item They obviously have the same EV
    \item Decoy 23\% \$23 may not work
    \item When competing options are objectively equivalent
  \end{itemize}
\end{flushleft}

\pause
\begin{flushleft}
  \textbf{Moving away from single choice DV}
  \begin{itemize}
    \item Elicit WTP?
    \item Ask for rankings?
    \item Allow multiple choices?
  \end{itemize}
\end{flushleft}


### Moving Forwards
\pause
\begin{flushleft}
  \textbf{What do you make of the results?}
\end{flushleft}
\begin{itemize}
  \item Risky decoy lowered "mistake" rate
  \item While safe decoy backfired
\end{itemize}

\pause
\begin{flushleft}
  \textbf{How to position this study?}
\end{flushleft}
\begin{itemize}
  \item Reference-dependent risk preference
  \item Regret theory — state-wise comparison
  \item Heuristics and biases — reduced mistakes
\end{itemize}

\pause
\begin{flushleft}
  \textbf{Next step?}
\end{flushleft}
\begin{itemize}
  \item Refine experimental design — a better control?
  \item More experimental or theoretical?
  \item Test in more realistic settings?
\end{itemize}

# Appendix
### Study 1: Frequency table
```{r}

study1_complete <- study1_complete %>%
  dplyr::mutate(
    condition = dplyr::recode_factor(
      condition,
      'risk_averse' = 'Safe',
      'risk_seeking' = 'Risky'
    ),
    condition = factor(condition, levels = c("Safe", "Risky")),  # ⬅ enforce order
    treatment = dplyr::recode_factor(
      treatment,
      'control' = 'Control',
      'treatment' = 'Treatment'
    ),
    treatment = factor(treatment, levels = c("Control", "Treatment")) # optional
  )
# Create table of counts
tab <- as.data.frame(table(
  Treatment = study1_complete$treatment,
  Choice    = study1_complete$raw_choice,
  Condition = study1_complete$condition
))

# Pivot wider
tab_wide <- tidyr::pivot_wider(
  tab,
  names_from = Choice,
  values_from = Freq
)

# ---- Convert to percentages ----
tab_wide <- tab_wide %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    total = sum(dplyr::c_across(where(is.numeric))),
    dplyr::across(where(is.numeric), ~ paste0(round(.x / total * 100, 1), "\\%"))
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(-total)# drop total if you don’t want it shown

tab_wide <- tab_wide %>%
  dplyr::mutate(
    risky = ifelse(
      risky %in% c(""), 
      kableExtra::cell_spec(risky, "latex", background = "yellow"), 
      as.character(risky)),
    decoy = ifelse(
      decoy %in% c(""), 
      kableExtra::cell_spec(decoy, "latex", background = "yellow"), 
      as.character(decoy)
    ))

library(kableExtra)
knitr::kable(
  tab_wide,
  format = "latex",
  escape = FALSE,   # ⬅ important so LaTeX code is kept
  booktabs = TRUE,
  col.names = c("Treatment", "Condition", "Decoy", "Risky", "Safe")
) %>%
  kable_styling(position = "center")
```

\vfill
\hfill \tiny *Full sample without excluding any participant

### Appendix: Switch to Focal, Conditioned on Not Choosing Focal in Control
```{r}
study1_change %>%
    dplyr::filter((control != "safe" & condition == "risk_averse")|
                  (control != "risky" & condition == "risk_seeking"))%>%
  dplyr::group_by(condition) %>%
  dplyr::summarize(mean = mean(change_from_any),
                   n = dplyr::n(),
                   sd = sd(change_from_any),
                   ci = qnorm(0.975) * sd/sqrt(n)) %>%
  ggplot2::ggplot(ggplot2::aes(x = condition,
                               y = mean,
                               fill = condition)) +
  ggplot2::geom_col(position = 'dodge') +
  ggplot2::geom_text(ggplot2::aes(x = condition,
                                  y = 0.05,
                                  label = paste0(round(mean*100, 2), "%", "\n",
                                                 "n = ", n)),
                     color = 'white',
                     lineheight = 0.75,
                     size = 2.5,
                     position = ggplot2::position_dodge(0.9)) +
  ggplot2::coord_cartesian(ylim = c(0, 1),
                           expand = 0) +
  ggplot2::scale_fill_manual(values = c('#998ec3', '#f1a340')) +
  ggplot2::geom_errorbar(ggplot2::aes(ymin = mean - ci,
                                      ymax = mean + ci),
                         position = ggplot2::position_dodge(0.9),
                         width = 0.2) +
  papaja::theme_apa() +
  ggplot2::theme(legend.position = c(1, 1.1),
                 legend.justification = c(1, 1),
                 legend.direction = "horizontal",
                 legend.background = ggplot2::element_rect(fill = "transparent", color = NA),
                 legend.text = ggplot2::element_text(size = 6),
                 legend.title = ggplot2::element_blank(),
                 axis.title.x = ggplot2::element_blank(),
                 axis.title.y = ggplot2::element_blank(),
                 axis.ticks.x = ggplot2::element_blank(),
                 strip.placement = "outside")+
  ggplot2::scale_y_continuous("Proportion Choosing Focal Lottery")

```

### Appendix: Move away from Focal, Conditioned on Already Choosing Focal in Control
```{r}
study1_change %>%
    dplyr::filter((control == "safe" & condition == "risk_averse")|
                  (control == "risky" & condition == "risk_seeking"))%>%
  dplyr::group_by(condition) %>%
  dplyr::summarize(mean = mean(move_away_oppo),
                   n = dplyr::n(),
                   sd = sd(move_away_oppo),
                   ci = qnorm(0.975) * sd/sqrt(n)) %>%
  ggplot2::ggplot(ggplot2::aes(x = condition,
                               y = mean,
                               fill = condition)) +
  ggplot2::geom_col(position = 'dodge') +
  ggplot2::geom_text(ggplot2::aes(x = condition,
                                  y = 0.05,
                                  label = paste0(round(mean*100, 2), "%", "\n",
                                                 "n = ", n)),
                     color = 'white',
                     lineheight = 0.75,
                     size = 2.5,
                     position = ggplot2::position_dodge(0.9)) +
  ggplot2::coord_cartesian(ylim = c(0, 1),
                           expand = 0) +
  ggplot2::scale_fill_manual(values = c('#998ec3', '#f1a340')) +
  ggplot2::geom_errorbar(ggplot2::aes(ymin = mean - ci,
                                      ymax = mean + ci),
                         position = ggplot2::position_dodge(0.9),
                         width = 0.2) +
  papaja::theme_apa() +
  ggplot2::theme(legend.position = c(1, 1.1),
                 legend.justification = c(1, 1),
                 legend.direction = "horizontal",
                 legend.background = ggplot2::element_rect(fill = "transparent", color = NA),
                 legend.text = ggplot2::element_text(size = 6),
                 legend.title = ggplot2::element_blank(),
                 axis.title.x = ggplot2::element_blank(),
                 axis.title.y = ggplot2::element_blank(),
                 axis.ticks.x = ggplot2::element_blank(),
                 strip.placement = "outside")+
  ggplot2::scale_y_continuous("Proportion Choosing Focal Lottery")

```

### Appendix: Heat Map -- Safe Condition

```{r}
study1_change_safe <- study1_change %>%
  dplyr::filter(condition == "risk_averse")%>%
  dplyr::mutate(
    control = dplyr::recode_factor(
      control,
      'safe' = 'Safe',
      'risky' = 'Risky',
      'decoy' = 'Decoy'),
    control = factor(control, levels = c("Decoy", "Risky", "Safe")),
    treatment = dplyr::recode_factor(
      treatment,
      'safe' = 'Safe',
      'risky' = 'Risky',
      'decoy' = 'Decoy'),
    treatment = factor(treatment, levels = c("Decoy", "Risky", "Safe")))

# Create table of counts
tab_safe <- as.data.frame(table(
  Control = study1_change_safe$control,
  Treatment = study1_change_safe$treatment))

# Pivot wider
tab_wide_safe <- tidyr::pivot_wider(
  tab_safe,
  names_from = Treatment,
  values_from = Freq)

# ---- Convert to percentages ----
tab_wide_safe <- tab_wide_safe %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    total = sum(dplyr::c_across(where(is.numeric))),
    dplyr::across(where(is.numeric), ~ paste0(round(.x / total * 100, 1), "\\%"))
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(-total)# drop total if you don’t want it shown



library(kableExtra)
knitr::kable(
  tab_wide_safe,
  format = "latex",
  escape = FALSE,   # ⬅ important so LaTeX code is kept
  booktabs = TRUE,
  col.names = c("Control", "Decoy", "Risky", "Safe"),
  caption = "Safe condition"
) %>%
  kable_styling(position = "center")



```

### Appendix: Heat Map -- Risky Condition

```{r}
study1_change_risky <- study1_change %>%
  dplyr::filter(condition == "risk_seeking")%>%
  dplyr::mutate(
    control = dplyr::recode_factor(
      control,
      'safe' = 'Safe',
      'risky' = 'Risky',
      'decoy' = 'Decoy'),
    control = factor(control, levels = c("Decoy", "Risky", "Safe")),
    treatment = dplyr::recode_factor(
      treatment,
      'safe' = 'Safe',
      'risky' = 'Risky',
      'decoy' = 'Decoy'),
    treatment = factor(treatment, levels = c("Decoy", "Risky", "Safe")))


# Create table of counts
tab_risky <- as.data.frame(table(
    Treatment = study1_change_risky$treatment,
  Control = study1_change_risky$control
))

# Pivot wider
tab_wide_risky <- tidyr::pivot_wider(
  tab_risky,
  names_from = Treatment,
  values_from = Freq)

# ---- Convert to percentages ----
tab_wide_risky <- tab_wide_risky %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    total = sum(dplyr::c_across(where(is.numeric))),
    dplyr::across(where(is.numeric), ~ paste0(round(.x / total * 100, 1), "\\%"))
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(-total)# drop total if you don’t want it shown


library(kableExtra)
knitr::kable(
  tab_wide_risky,
  format = "latex",
  escape = FALSE,   # ⬅ important so LaTeX code is kept
  booktabs = TRUE,
  col.names = c("Control", "Decoy", "Risky", "Safe"),
  caption = "Risky condition"
) %>%
  kable_styling(position = "center")

```

